The core transcriptome of the mammal placenta

### 2016-08-16 manuscript/manuscript.Rnw
plot ancestral trees

<<plot.defaults,echo=FALSE,results="hide">>=
grad.def <- list(low = "#132B43", high = "#56B1F7", space = "Lab", na.value = NA)

@@

\begin{figure*}
<<psg.ctsl.evolution.tree,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=
plot.label.size <- 3
axis.text.size <- 3.5

pushViewport(viewport(layout=grid.layout(nrow=2,ncol=2),clip="off"))
tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PSG1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,min.fpkm=1,
                 vp=viewport(layout.pos.col=1,layout.pos.row=1,clip="off"),
                 axis.text.size=axis.text.size,
                 fontsize=plot.label.size,
                 )

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CTSL" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(gene.tree,gene.tree.table,gene.tree.expression,min.fpkm=1,
                 subtree=c("ENSMODG00000002844","ENSMUSG00000055679"),
                 vp=viewport(layout.pos.col=2,layout.pos.row=1,clip="off"),
                 fontsize=4,
                 width=3,
                 axis.text.size=axis.text.size,
                 offset.ratio=1
                 )

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PRSS16" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 vp=viewport(layout.pos.col=1,layout.pos.row=2,clip="off"),
                 fontsize=4,
                 width=3,
                 axis.text.size=axis.text.size,
                 offset.ratio=1.2
                 )

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PAG2" & species=="bos taurus",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,min.fpkm=1,
                 subtree=c("ENSSSCG00000006792",
                           "ENSBTAG00000046085"),
                 vp=viewport(layout.pos.col=2,layout.pos.row=2,clip="off"),
                 axis.text.size=axis.text.size,
                 fontsize=plot.label.size,
                 offset.ratio=0.6
                 )

popViewport(1)
grid.text(c("A","B","C","D"),
          x=unit(c(0,0.5,0,0.5),"npc"),y=unit(c(1,1,0.5,0.5),"npc"),
          just=c("left","top"),
          gp=gpar(fontsize="48")
          )
@ 
\caption{\genetreeexplanation{(A) \textit{PSG1}, (B) \textit{CTSL}, (C) \textit{PRSS16}, and (D) \textit{PAG2}}}
\label{fig:pag2_prss16_ctsl_psg_evolution_tree}
\end{figure*}

@@

\begin{figure}
<<hormone.genes.plot,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=

@@

\begin{figure}
<<chorionic.somatomammotropin.alpha.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

plot.label.size <- 4.5
pushViewport(viewport(layout=grid.layout(nrow=1,ncol=2,widths=c(2,3))))

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CGA" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 vp=viewport(layout.pos.col=1,layout.pos.row=1),
                 offset.ratio=0.7,
                 fontsize=plot.label.size,
                 axis.text.size=3.5
                 )

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CGB1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 vp=viewport(layout.pos.col=2,layout.pos.row=1),
                 subtree=c("ENSMODG00000013768","ENSG00000104818"),
                 offset.ratio=1,
                 fontsize=plot.label.size,
                 axis.text.size=3.5
                 )
popViewport(1)
grid.text(c("A","B"),
          x=unit(c(0,2/5),"npc"),y=unit(c(1,1),"npc"),
          just=c("left","top"),
          gp=gpar(fontsize="48")
          )
@ 
\caption{\genetreeexplanation{(A) \textit{CGA} (B) \textit{CGB}}}
\label{fig:cga_cgb_evolution_tree}
\end{figure}

>>>>

<<load_plot_ancestral_trees,message=FALSE,echo=FALSE,results="hide",cache.extra=file.info(paste0("../common_r_code/",c("plot_ancestral_state_tree.R")))[,"mtime"]>>=
for (a in paste0("../common_r_code/",c("plot_ancestral_state_tree.R"))) {
    source(a)
}
@

<<plot.defaults,echo=FALSE,results="hide">>=
#grad.def <- list(low = "#132B43", high = "#56B1F7", space = "Lab", na.value = NA)
grad.def <- list(low = "gray90", high = "blue", space = "Lab", na.value = NA)

@@

\begin{figure*}[htb]
<<psg.ctsl.evolution.tree,echo=FALSE,out.width="\\textwidth",fig.width=8,fig.height=8>>=
pushViewport(viewport(layout=grid.layout(nrow=2,ncol=2),clip="off"))
print(plot.ancestral.state.tree(gene_selector=quote(gene_short_name=="PSG1" & species=="homo sapiens"),
                                genes.to.tree=genes.to.tree,
                                combined.fpkm=combined.fpkm,
                                subtree=c("ENSG00000221826","ENSG00000124469"),
                                trees=trees
                                )+
      geom_tiplab(size=2,color="black",fontface="italic")+xlim(0,0.2)+
      theme(legend.position="none"),
      vp=viewport(layout.pos.col=1,layout.pos.row=1,clip="off"))
print(plot.ancestral.state.tree(gene_selector=quote(gene_short_name=="CTSL" & species=="homo sapiens"),
                                genes.to.tree=genes.to.tree,
                                combined.fpkm=combined.fpkm,
                                subtree=c("ENSMODG00000002844","ENSMUSG00000055679"),
                                trees=trees)+
      geom_tiplab(size=2,color="black",fontface="italic")+xlim(0,1)+
      theme(legend.position="none"),
      vp=viewport(layout.pos.col=2,layout.pos.row=1,clip="off"))
print(plot.ancestral.state.tree(gene_selector=quote(gene_short_name=="PRSS16" &
                                                    species=="homo sapiens"),
                                genes.to.tree=genes.to.tree,
                                subtree=c(   "ENSG00000112812",
                                          "ENSLAFG00000032211"),
                                combined.fpkm=combined.fpkm,
                                trees=trees)+
      geom_tiplab(size=2,color="black",fontface="italic")+xlim(0,1.2)+
      theme(legend.position="none"),
      vp=viewport(layout.pos.col=1,layout.pos.row=2,clip="off"))
print(plot.ancestral.state.tree(gene_selector=quote(gene_short_name=="PAG2" &
                                                    species=="bos taurus"),
                                genes.to.tree=genes.to.tree,
                                combined.fpkm=combined.fpkm,
                                subtree=c("ENSSSCG00000013097",
                                          "ENSBTAG00000046085"),
                                trees=trees) +
      geom_tiplab(size=2,color="black",fontface="italic") +
      xlim(0,0.8),
      vp=viewport(layout.pos.col=2,layout.pos.row=2,clip="off"))
popViewport(1)

grid.text(c("A","B","C","D"),
          x=unit(c(0,0.5,0,0.5),"npc"),y=unit(c(1,1,0.5,0.5),"npc"),
          just=c("left","top"),
          gp=gpar(fontsize="24")
          )
@ 
\caption{\genetreeexplanation{(A) \textit{PSG1}, (B) \textit{CTSL}, (C) \textit{PRSS16}, and (D) \textit{PAG2}}}
\label{fig:pag2_prss16_ctsl_psg_evolution_tree}
\end{figure*}

@@

\begin{figure}[htb]
<<hormone.genes.plot,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=

@@

\begin{figure}[htb]
<<chorionic.somatomammotropin.alpha.evolution.tree,echo=FALSE,fig.width=4,fig.height=4,warning=FALSE,message=FALSE>>=

pushViewport(viewport(layout=grid.layout(nrow=1,ncol=2,widths=c(2,3))))

print(plot.ancestral.state.tree(gene_selector=quote(gene_short_name=="CGA" &
                                                    species=="homo sapiens"),
                                genes.to.tree=genes.to.tree,
                                combined.fpkm=combined.fpkm,
                                trees=trees
                                )+
      geom_tiplab(size=2,color="black",fontface="italic")+xlim(0,0.8)+
      theme(legend.position="none"),
      vp=viewport(layout.pos.col=1,layout.pos.row=1,clip="off"))
print(plot.ancestral.state.tree(gene_selector=quote(gene_short_name=="CGB1" &
                                                    species=="homo sapiens"),
                                genes.to.tree=genes.to.tree,
                                subtree=c("ENSMODG00000013768","ENSG00000104818"),
                                combined.fpkm=combined.fpkm,
                                trees=trees
                                )+
      geom_tiplab(size=2,color="black",fontface="italic")+xlim(0,1),
      vp=viewport(layout.pos.col=2,layout.pos.row=1,clip="off"))
popViewport(1)
grid.text(c("A","B"),
          x=unit(c(0,2/5),"npc"),y=unit(c(1,1),"npc"),
          just=c("left","top"),
          gp=gpar(fontsize="24")
          )
@ 
\caption{\genetreeexplanation{(A) \textit{CGA} (B) \textit{CGB}}}
\label{fig:cga_cgb_evolution_tree}
\end{figure}
